Constraints on Assembly Bias from Galaxy Clustering.

### 2016-06-07 ms.tex
more updates to intro and bib-file

\newcommand{\mvir}{M_{\rm vir}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\ncen}{N_{\rm cen}}
\newcommand{\nsat}{N_{\rm sat}}
\newcommand{\ngal}{N_{\rm gal}}
\newcommand{\abias}{\mathcal{A}_{\rm bias}}

@@

\newcommand{\wprp}{w_{\mathrm{p}}}
\newcommand{\rp}{r_{\mathrm{p}}}
\newcommand{\magr}{M_r}

\newcommand{\bit}{\begin{itemize}}
\newcommand{\eit}{\end{itemize}}
\newcommand{\ben}{\begin{enumerate}}
\newcommand{\een}{\end{enumerate}}

@@

\title[Clustering Constraints on Assembly Bias]
{
Constraints on Assembly Bias from Galaxy Clustering
}

@@

\author[Zentner et al.]
{Andrew R. Zentner$^{1}$, Andrew P. Hearin$^{2}$, Frank C. van den Bosch$^{3},$ \newauthor
Antonio Villareal$^{1}$, Johannes Lange$^{3}$, P. Rogers Nelson$^{4}$, Probably some others of unspecified origin\\ \\
$^1$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260\\
$^2$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
$^3$Department of Astronomy, Yale University, P.O. Box 208101, New Haven, CT\\
$^4$Paisley Park, Chanhassen, MN\\
}

@@

\maketitle
%----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%  A B S T R A C T %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ARZ: Important points:
% -supersedes previous analyses even for standard HOD.
% -cannot rule out assembly bias using galaxy clustering.
% -some samples favor assembly bias.
\begin{abstract}
We constrain the newly-introduced Decorated HOD model using SDSS DR7 measurements of projected galaxy clustering, $\wprp,$ and number density, $n_{\rm d},$ made from r-band luminosity-threshold samples. The Decorated HOD is a model for the galaxy--halo connection that augments the traditional Halo Occupation Distribution (HOD) by allowing for the possibility of {\em assembly bias}: galaxy luminosity may be correlated with dark matter halo properties besides virial mass $\mvir$ alone. We demonstrate that it is not possible to rule out assembly bias using DR7 measurements of $\wprp$ and $n_{\rm d}.$ Moreover, galaxy samples $\magr<-20, -20.5$ favor strong levels of central galaxy assembly bias: high-concentration halos are more likely to host a central galaxy relative to low-concentration halos of the same $\mvir.$ We rule out zero assembly bias with high significance for these samples. Satellite assembly bias only becomes significant for the faintest sample we study, $\magr<-19.$ We find no evidence for assembly bias in the $\magr<-21$ sample. Because the Decorated HOD subsumes the traditional HOD, and additionally because of errors made in the original analysis of these samples, our results differ from and supersede existing HOD constraints from low-redshift clustering measurements. 
\end{abstract}

@@

Halo occupation modeling has frequently 
been used to interpret large-scale structure measurements and exploit these measurements 
to constrain galaxy formation models and cosmology \citep[e.g.,][]{yang03,tinker05,zehavi05a,
porciani06,vdBosch07,Zheng07,conroy_wechsler09,yang09b,zehavi_etal11,guo_etal11b,
wake_etal11,yang11a,yang12,leauthaud_etal12,tinker_etal13,cacciato_etal13,
more_etal13,guo_etal14,zu_mandelbaum15b}. The key assumptions underlying 
halo occupation modeling are: (1) all galaxies reside in dark matter halos that are biased 
tracers of the density field; and (2) galaxies occupy halos as a function of halo mass 
only. It is now well known that halo bias depends upon halo properties other than mass
\citep[e.g.][]{gao_etal05,wechsler06,gao_white07,zentner07,dalal_etal08,lacerna11}, 
an effect called {\em halo assembly bias}. If galaxies occupy halos as a function of properties 
that affect halo clustering, other than mass, then standard halo occupation methods will be subject to a 
systematic error due to {\em galaxy assembly bias}. Several of us have previously shown that 
this error can be significant in an analysis of galaxy clustering and can bias inferences 
about many aspects of galaxy evolution \citep{zentner_etal14}. Consequently, we have 
developed halo occupation models that enable galaxies to occupy halos in a manner 
that depends upon several halo properties \citep{hearin_etal16}. In this paper, we 
revisit the interpretation of luminosity-dependent galaxy clustering in the Sloan Digital Sky 
Survey (SDSS) Data Release 7 (DR7) data, analyzed previously by \citet{zehavi_etal11}, in the context 
both standard halo occupation models and these new models.

Some history here, blah blah blah, Frank loves this stuff. Therefore, I'm sure this part of the introduction 
will get quite lengthy.

Our work is important for several reasons. Our work is a re-analysis of the SDSS DR7 data 
the overcomes shortcomings of previous analyses. In particular, we use direct population of galaxies 
in a cosmological simulation so that delicate issues present in analytic modeling, such as scale-dependent 
halo bias, are treated exactly.  The simulation we use is based on the latest Planck cosmological parameters, 
updating previous work. Furthermore, many important differences between this work and the previous 
work of \citet{zehavi_etal11} were caused by insufficiently large Monte Carlo Markov Chain (MCMC) sampling 
by \citet[][Z. Zheng \& I. Zehavi, Private Communication]{zehavi_etal11}. Therefore, our analysis 
{\em supersedes previous analyses even in the case of standard halo occupation models}.

Furthermore, our work demonstrates explicitly that significant assembly bias in 
$M_r$-selected samples from SDSS DR7 cannot be ruled out based on a standard 
analysis of galaxy clustering alone. In fact, we find that the clustering of galaxies in 
several luminosity threshold samples 
{\em favors galaxy assembly bias to a degree that is statistically significant}. 
As suggested by the previous work of \citet{zentner_etal14}, 
this conclusion could have important consequences for the 
interpretation of both extant and forthcoming galaxy survey data.

@@

To generate predictions for galaxy clustering, we directly populate dark matter halos with mock galaxies using {\tt Halotools}. 
We explore halo occupation distribution models (HOD) models in this work 
\citep[e.g.][]{seljak00,ma_fry00,scoccimarro01a,berlind02}, though other 
techniques that can be used to interpret such data, such as the conditional luminosity function 
\citep[CLF, e.g.,][]{yang03,vdBosch13}, exist. In this section, we review the ``standard" HOD 
model used in the present work which assumes that there is no galaxy assembly bias. We refer to 
such a model as ``standard" because all HOD analyses of galaxy clustering to date have assumed no 
galaxy assembly bias. In the following section, we describe 
the Decorated HOD model described in \citet{hearin_etal16}. In both cases, we will only 
review the salient features of our methodology briefly; 
interested readers can always refer to {\tt halotools.readthedocs.io} and 
\citet{hearin_etal16} for further details.

\subsubsection{Simulation}

All of our analyses are based on the Bolshioi-Planck simulation \citep{riebe_etal11}. 
Bolshoi-Planck was run with cosmological parameters based on \citet{planck13}, $\Omega_{\Lambda} = 0.693, \Omega_{\rm m} = 0.307, \Omega_{\rm b} = 0.048, h = 0.7, {\rm n_s}=0.96$ and $\sigma_8 = 0.82,$ within a cubic box 
$250$ ${\rm Mpc}/h$ on a side, requiring a particle mass of $m_{\rm p} = 1.35\times10^{8}M_{\odot}/h.$ 
Further information about the Bolshoi-Planck simulation is available at \url{https://www.cosmosim.org}.

@@

In standard HOD models, central galaxies and satellite galaxies are treated separately, so 
the model is specified by two probability distributions, one for each type of galaxy. 
The galaxy-halo connection is specified in terms of $P(\ncen|\mvir)$ and $P(\nsat|\mvir),$ 
the probability that a halo of mass $\mvir$ hosts $\ncen$ central and $\nsat$ satellite galaxies, 
respectively. $P(\ncen|\mvir)$ is typically a nearest-integer distribution, as a host halo has only 
either zero or one central galaxy. Consequently, the occupation statistics of central galaxies are 
specified by the first moment of $P(\ncen|\mvir)$, which we model as 
%
\begin{equation}
\label{eq:ncen}
\langle N_{\mathrm{cen}} \rangle(\mvir) =
        \frac{1}{2}\left( 1 +
        \mathrm{erf}\left( \frac{\log (\mvir) -
        \log (M_{\rm min})}{\sigma_{\log M}} \right) \right)
\end{equation}
%
For every host halo in the catalog we draw a random number from a uniform distribution $\mathcal{U}(0, 1);$ for a host halo of present-day virial mass $\mvir,$ a central galaxy is assigned to the halo if the associated random number is less than $\langle N_{\mathrm{cen}} \rangle(\mvir);$ halos with random values exceeding  $\langle N_{\mathrm{cen}} \rangle(\mvir)$ are left devoid of centrals. 
The parameter $\log (M_{\rm min})$ specifies the halo mass at which the halo has a 50\% probability of hosting a 
central galaxy, while the parameter $\sigma_{\log M}$ specifies the rate at which $\langle N_{\mathrm{cen}} \rangle (\mvir)$ 
transitions from zero to unity, with smaller values of $\sigma_{\log M}$ corresponding to a more rapid transition.

We model the distribution $P(\nsat|\mvir)$ as a Poisson distribution with first moment given by a power law, 
%
\begin{equation}
\label{eq:nsat}
\langle N_{\rm sat}\rangle(\mvir) = \left( \frac{\mvir - M_0}{M_1} \right)^{\alpha}. 
\end{equation}
%
The parameter $M_0$ allows the power-law to be truncated more rapidly at low masses and we 
set $\langle N_{\rm sat}\rangle (\mvir) = 0$ for $\mvir < M_0$.

The five parameters of the standard HOD models are varied in standard analyses are $\log (M_{\rm min})$, 
$\sigma_{\log M}$, $\alpha$, $\log (M_1)$, and $\log (M_0)$, though, as we show below, central galaxies usually 
outnumber satellite galaxies in by a factor of several, so $\log (M_{\rm min})$ and $\sigma_{\log M}$ usually 
vary along a narrow degeneracy that fixes the total galaxy number density to the observed value. There are 
many particular choices that can be made for the functional forms of $\langle N_{\rm cen}\rangle (\mvir)$ and 
$\langle N_{\rm sat}\rangle (\mvir)$. We have made choices that mimic the standard SDSS DR7 analysis of 
\citet{zehavi_etal11}, to expedite comparisons with their results.

@@

We model the radial velocity distribution of satellite galaxies as a Gaussian with 
first moment equal to the host halo velocity and second moment equal to the 
solution to the isotropic Jeans equation for an NFW profile \citep{more09b}, 
\begin{equation}
\sigma^{2}_{r}(\tilde{r}|c) = V_{\rm vir}^{2}\frac{c^{2}\tilde{r}(1 + c\tilde{r})^{2}}{g(c)}\int_{c\tilde{r}}^{\infty}{\rm d}y\frac{g(y)}{y^{3}(1 + y)^{2}}, 
\end{equation}
where $\tilde{r} = r/\rvir,$ $g(x) = \rm{ln}(1+x) - x / (1+x),$ and $V_{\rm vir}^{2} = G\mvir/\rvir.$ 
We assume that velocities are isotropic, setting peculiar velocities in the simulation $z$-dimension  
equal to random draws from the above distribution. We place mock galaxies 
into redshift-space by replacing each galaxy's $z$-coordinate with $z \rightarrow z + V_{\rm z}/H_0.$

@@

In addition to the standard occupation statistics described in the previous section, 
in this paper we also use the decorated HOD formalism to connect galaxies t
o dark matter halos in a manner that has simultaneous dependence on both $\mvir$ and halo concentration. 
Briefly, we use Equations \ref{eq:ncen} and \ref{eq:nsat} as our ``baseline" first occupation moments. 
At fixed $\mvir,$ halos are divided into one of two categories, those of high- and low-concentration, 
depending on whether the concentration of the halo places it above or below the rank-order percentile 
$f_{\rm split},$ which we keep fixed to $f_{\rm split}=0.5$ throughout the paper for simplicity. 
High-concentration halos have a different first occupation moment relative to low-concentration 
halos of the same mass, $\langle\ngal|\mvir, c_{\rm high}\rangle \neq \langle\ngal|\mvir, c_{\rm low}\rangle.$ The difference between the first moment of high- and low-concentration halos is modulated by $\abias,$ the novel parameter of the decorated HOD governing assembly bias. Values of $\abias=\pm1$ correspond to the maximum strength of assembly bias allowable by the constraint that the model preserves the marginalized first moment, $\langle\ngal|\mvir\rangle;$ thus regardless of the value of $\abias,$ in the decorated HOD the marginalized first moment of centrals and satellites are {\em unchanged from the baseline value defined by Equations \ref{eq:ncen} and \ref{eq:nsat}}. 
Decorated HOD models all have the same HODs, when averaged over all halos at fixed mass, as standard HOD models. The only 
change in Decorated HOD models is whether or not an additional property also modulates halo occupation at fixed halo mass. A 
value of $\abias = 0$ indicates no galaxy assembly bias whatsoever. We refer the reader to \citet{hearin_etal16} for 
further details about the decorated HOD.

In the present work, we fix our model to the simplest class of galaxy assembly bias models, though a recipe for 
generalizing to more complicated models can be found in \citet{hearin_etal16}. In particular, we split halos into 
two populations as specified in the previous paragraph. We then populate halos with satellite galaxies specified 
by an assembly bias parameter $-1 \le A_{\rm sat} \le 1$ and central galaxies with a distinct assembly bias 
parameter $-1 \le A_{\rm cen} \le 1$. These two additional parameters are varied in our analyses that include 
assembly bias, making the number of parameters that vary in these analyses seven. As we show below, 
these additional parameters are in many instances poorly constrained by clustering data of the quality of 
SDSS DR7 alone, so exploring more complex models of assembly bias does not yet seem justified.

@@

property that we assume to determine the galaxy HOD is halo concentration. $A_{\rm cen,sat}=0$ means that there is no 
assembly bias while $A_{\rm cen,sat}=1$ ($A_{\rm cen,sat}=-1$) means that galaxy abundance is maximally 
correlated (anticorrelated) with halo concentration at fixed $\mvir.$ Thus the $A_{\rm cen,sat}$ parameters span the range $[-1, 1].$ 
If the constraints on $A_{\rm cen}$ and $A_{\rm sat}$ are unspecified in the table, then the model used to interpret the data 

@@

A second noteworthy difference between the present work and that of \citet{zehavi_etal11} is that we 
find many parameters to be notably more poorly constrained. At the lower luminosity thresholds, for example, 
we constrain $\log (M_{\rm min})$ and $\sigma_{\log M}$ with several times lower precision than 
\citet{zehavi_etal11}. We do not show our constraints on $\log (M_0)$ as they are very poor in all cases, 
with 1-sigma constraints on the order of $\pm \ge 1$~dex for nearly all samples. In several cases, the constraint on 
$\log (M_0)$ is determined by the prior given in Table~\ref{table:priors}. This is in stark contrast to several of the 
results of \citet{zehavi_etal11}. For example, for the threshold sample with $M_r < -19.5$ ($M_r < -20.5$), 
\citet{zehavi_etal11} quote $\log (M_0) = 12.23 \pm 0.17$ ($12.35 \pm 0.24$), whereas we infer 
$\log (M_0) = 11.38^{+0.95}_{-1.57}$ ($11.19^{+0.89}_{-1.39}$). Examining the form of Eq.~(\ref{eq:nsat}), 
it is sensible that the parameter $\log (M_0)$ should be unconstrained at the lower end, because the 
value of $M_0$ does not alter the predicted satellite number once $M_0 \ll M_1$. 
Therefore, on these intuitive grounds alone it seems likely that the tighter constraints quoted by \citet{zehavi_etal11} must be an error.

Additionally, we have confirmed with a subset of the authors of \citet{zehavi_etal11} 
that the number of MCMC samples they included in their analysis was 
insufficient in a number of cases and that this can lead to a significant underestimation of the uncertainties 
on the inferred parameters, especially $\log (M_{\rm min})$ and $\sigma_{\log M}$ 
(Z. Zheng \& I. Zehavi, private communication). The analysis of \citet{zehavi_etal11} 
used $10^4$ samples, whereas we find several $\times 10^6$ samples are often necessary for convergence. 
Additionally, we have recreated qualitatively similar behavior by considering only small subsets of our 
full MCMC chains. Consequently, insufficient sampling of the posterior seems to be the likely resolution 
of the discrepancies between our work and that of \citet{zehavi_etal11}.

Two degeneracies are manifest in Fig.~\ref{fig:Mr20triangle} that are common to all of our 
analyses. The parameters $\log (M_1)$ and $\alpha$ are degenerate with each other and positively 
correlated. The parameter $M_1$ is the mass scale at which a halo has one satellite on 
average, and $\alpha$ is the power-law index describing the dependence of average satellite 
number on halo mass. Increasing $M_1$ {\em decreases} the number of satellites in massive halos 
by increasing the mass scale where the power law abundance becomes operative. An increase in $\alpha$ can partly 
compensate for an increase in $M_1$ by increasing the rate at which average satellite number 
grows with halo mass.

@@

We have re-analyzed the SDSS DR7 measurements of projected galaxy clustering,  $\wprp,$ and number density, $n_{\rm g},$ originally published in \citet{zehavi_etal11}. Our work is especially novel in that we provide the first quantitative constraints on assembly bias derived from the Decorated HOD, an extension to the traditional HOD introduced in \citet{hearin_etal16} developed for exactly this purpose. 
We enumerate our most important conclusions below. 
\ben
\item It is not possible to rule out galaxy assembly bias using SDSS DR7 measurements of $\wprp$ and $n_{\rm g}.$
\item Decorated HOD fits to $\wprp$ and $n_{\rm g}$ favor significant levels of assembly bias, particularly in the lower luminosity thresholds we sample. 
\een
As $\wprp$ and $n_{\rm g}$ are already very well-measured in DR7, it is likely that further improvements on assembly bias constraints at low-redshift will require additional observational measurements, e.g., galaxy--galaxy lensing, group statistics, etc.

>>>>

\newcommand{\mvir}{M_{\rm vir}}
\newcommand{\vmax}{v_{\rm max}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\ncen}{N_{\rm cen}}
\newcommand{\nsat}{N_{\rm sat}}
\newcommand{\ngal}{N_{\rm gal}}
\newcommand{\abias}{\mathcal{A}_{\rm bias}}

%%%%% less than approximate and all that

\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\prosima{$\; \buildrel \propto \over \sim \;$}
\def\gsim{\lower.7ex\hbox{\gtsima}}
\def\lsim{\lower.7ex\hbox{\ltsima}}
\def\simgt{\lower.7ex\hbox{\gtsima}}
\def\simlt{\lower.7ex\hbox{\ltsima}}
\def\simpr{\lower.7ex\hbox{\prosima}}
\def\la{\lsim}
\def\ga{\gsim}
\def\lta{\la}
\def\gta{\ga}

@@

\newcommand{\wprp}{w_{\mathrm{p}}}
\newcommand{\rp}{r_{\mathrm{p}}}

@@

\title[Clustering Constraints on Assembly Bias]
{
Constraints on Assembly Bias from Galaxy Clustering. 
}

@@

\author[Zentner et al.]
{Andrew R. Zentner$^{1}$, Andrew P. Hearin$^{2}$, Frank C. van den Bosch$^{3},$ \newauthor
Antonio Villareal$^{1}$, Johannes U. Lange$^{3}$, P. Rogers Nelson$^{4}$\\ \\
$^1$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260\\
$^2$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
$^3$Department of Astronomy, Yale University, P.O. Box 208101, New Haven, CT\\
$^4$Paisley Park, Chanhassen, MN\\
}

@@

\maketitle
%----------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%  A B S T R A C T %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

We fit SDSS DR7 data with models that include assembly bias.

Important points:
-supersedes previous analyses even for standard HOD.
-cannot rule out assembly bias using galaxy clustering.
-some samples favor assembly bias.

\end{abstract}

@@

For more than a decade, halo occupation modeling has been used to
interpret large-scale structure measurements and exploit these
measurements to constrain galaxy formation models and cosmology
\citep[e.g.,][]{yang03,tinker05,zehavi05a,
  porciani06,vdBosch07,Zheng07,conroy_wechsler09,yang09b,zehavi_etal11,guo_etal11b,
  wake_etal11,yang11a,yang12,leauthaud_etal12,rod_puebla12,tinker_etal13,cacciato_etal13,
  more_etal13,guo_etal14,zu_mandelbaum15b}. The key assumptions
underlying halo occupation modeling are: (1) all galaxies reside in
dark matter halos that are biased tracers of the density field; and
(2) galaxies occupy halos as a function of halo masses only. It is now
well known that halo bias depends upon halo properties other than mass
\citep[e.g.][]{gao_etal05,wechsler06,gao_white07,zentner07,dalal_etal08,lacerna11},
an effect called halo assembly bias. If galaxies occupy halos as a
function of properties other than halo mass, then standard halo
occupation methods will be subject to a systematic error due to galaxy
assembly bias. Several of us have previously shown that this error can
be significant in an analysis of galaxy clustering and can bias
inferences about many aspects of galaxy evolution
\citep{zentner_etal14}. Consequently, we have developed halo
occupation models that enable galaxies to occupy halos in a manner
that depends upon several halo properties \citep{hearin_etal16}.  See
also \citet{Yao_etalXX} for a similar approach. In this paper, we
revisit the interpretation of luminosity-dependent galaxy clustering
in the Sloan Digital Sky Survey (SDSS) Data Release 7 (DR7) data,
analyzed previously by \citet{zehavi_etal11}, in the context both
standard halo occupation models and these new models.

The ultimate goal of halo occupation modeling is to characterize the
galaxy-dark matter connection by constraining the probability
distribution $P(X_{\rm g}|Y_{\rm h})$. Here $X_{\rm g} = {x_1, x_2,
  ...x_n}$ is a set of $n$ galaxy properties, and $Y_{\rm h} = {y_1,
  y_2,...,y_m}$ a set of $m$ halo properties. Accurate knowledge of
this probability distribution yields invaluable insight to the
processes of galaxy formation, and allows one to make accurate
predictions for the clustering of galaxies as function of any property
$x_i$. Most halo occupation studies to date have restricted themselves
to only one halo property, namely halo mass.  This includes all Halo
Occupation Distribution (HOD) models, which aim to constrain
$P(N|\mvir)$, the probability that a halo of mass $M$ contains $N$
galaxies of some specified properties
\citep[e.g.,][]{kauffmann_etal97,jing98,benson01,berlind02}, and all
Conditional Luminosity Function (CLF) models, which aim to constrain
$\Phi(L|M)$, the luminosity (or stellar mass) function of galaxies in
haloes of mass $M$ \citep[e.g.,][]{yang03,vdBosch03a,cooray06}. The
HOD and CLF models are mainly used in combination with the halo model,
which describes the non-linear matter field in terms of halo building
blocks \citep[e.g.,][]{seljak00,cooray02}, to predict the clustering
of galaxies as function of luminosity or stellar mass.  Although
entirely analytical, and therefore fast, the downside of this method
is that it is extremely challenging to construct a model that is
accurate at the few percent level. The main issues are halo exclusion
\citep[see][]{smith_etal07,vdBosch13}, scale dependence of the
halo bias \citep[see][]{tinker05}, and the halo assembly bias mentioned
above.

These issues are trivially avoided by directly populating dark matter
haloes in numerical simulations with galaxies. This can be done using
either a parameterized model for the halo occupation statistics
\citep[e.g.,][]{Yang_etal04,moster10;behroozi10,behroozi13} or by
rank-order matching galaxy properties (i.e., luminosity or stellar
mass) to halo properties (i.e., halo mass). The latter method is known
as subhalo abundance matching (SHAM).  Although many SHAM studies use
halo mass as the halo property of choice
\citep[e.g.,][]{vale_ostriker04,vale_ostriker06,conroy07,conroy_wechsler09},
a number of studies have used alternative halo properties, such as the
maximum circular velocity, $\vmax$, or its peak value over the halo's
history, $V_{\rm peak}$
\citep[e.g.][]{tasitsiomi_etal04,conroy06,marin_etal08,reddick_etal13}

Since, at fixed halo mass, $\vmax$ is correlated with halo
concentration, and thus with halo assembly time, the resulting
occupation models have assembly bias build in
\citep[see][]{Zentner_etal14}; i.e., at fixed halo mass, galaxy
properties will be correlated with halo assembly
time. \citet{lehmann_etal15} took this one step further, and explored
SHAM models in which the halo property varies between $M$ and $\vmax$
via a continuously-valued parameter. Using SDSS clustering
measurements to constrain their model, they conclude that galaxy
luminosity reveals a significant dependence on halo
concentration. Since the latter is strongly correlated with halo
formation time \citep[e.g.,][]{wechsler02,zhao_etal09,ludlow_etal13},
their conclusions argue for non-zero assembly bias in the galaxy
population.

In addition to these studies that have focussed on constraining the
relation between a {\it primary} galaxy property (e.g., luminosity or
stellar mass) and a {\it primary} halo property (e.g., $M$ or
$\vmax$), there have also been a number of studies that relate
{\it secondary} galaxy properties (e.g., color or specific star
formation rate) to a {\it secondary} halo property (e.g., halo
formation time or concentration). The archetype of these model is the
`age-matching' model developed by \citet{HW13a}, in which galaxy color
is rank-order matched, at fixed stellar mass, to some quantity related
to the assembly history of the galaxy's (sub)halo.
Paranjape \etal 2015...

Another approach that 
This was taken one step further by Hearin ccc, who developed an
extension of SHAM, called age-matching. In age-matching a

Our work is important for several reasons. Our work is a re-analysis
of the SDSS DR7 data that overcomes shortcomings of previous
analyses. In particular, we use direct population of galaxies in a
cosmological simulation so that delicate issues present in analytic
modeling, such as scale-dependent halo bias, are treated exactly.  The
simulation we use is based on the latest Planck cosmological
parameters, updating previous work. Furthermore, many important
differences between this work and the previous analysis by
\citet{zehavi_etal11} result from the fact that the Monte Carlo Markov
Chains used by Zehavi \etal where too small and had therefore not
converged (Z. Zheng \& I. Zehavi, Private Communication).  Therefore,
our analysis {\em supersedes previous analyses even in the case of
  standard halo occupation models}.

Furthermore, our work demonstrates explicitly that significant
assembly bias in $M_r$-selected samples from SDSS DR7 cannot be ruled
out based on a standard analysis of galaxy clustering only. In fact,
in agreement with \citet{lehmann_etal15}, we find that several samples
{\em favor galaxy assembly bias to a degree that is statistically
  significant}. As demonstrated by \citet{zentner_etal14}, this
conclusion could have important consequences for the interpretation of
both extant and forthcoming data.

@@

To generate predictions for galaxy clustering, we directly populate dark matter halos with mock galaxies using {\tt Halotools}. In this section we review the ``standard" HOD-style model used in the present work, and in the following section we use the Decorated HOD model described in \citet{hearin_etal16}. In both cases, we will only briefly review the salient features of our methodology here; interested readers can always refer to {\tt halotools.readthedocs.io} for further details.

\subsubsection{Simulation}
All of our analyses are based on the Bolshioi-Planck simulation \citep{riebe_etal11}. 
Bolshoi-Planck was run with cosmological parameters based on \citet{planck13}, $\Omega_{\Lambda} = 0.693, \Omega_{\rm m} = 0.307, \Omega_{\rm b} = 0.048, h = 0.7, {\rm n_s}=0.96$ and $\sigma_8 = 0.82,$ and with a box size of $250$ ${\rm Mpc}/h$ and a particle mass of $m_{\rm p} = 1.35\times10^{8}M_{\odot}/h.$ Further information about the Bolshoi-Planck simulation is available at \url{https://www.cosmosim.org}.

@@

In standard HOD models, there are two types of galaxies: centrals and satellites. The galaxy-halo connection is specified in terms of $P(\ncen|\mvir)$ and $P(\nsat|\mvir),$ the probability that a halo of mass $\mvir$ hosts $\ncen$ central and $\nsat$ satellite galaxies, respectively. $P(\ncen|\mvir)$ is typically a nearest-integer distribution, as a host halo has only either zero or one central galaxy. So occupation statistics of central galaxies are specified by the first moment of $P(\ncen|\mvir)$, which we model as

\begin{equation}
\label{eq:ncen}
\langle N_{\mathrm{cen}} \rangle(\mvir) =
        \frac{1}{2}\left( 1 +
        \mathrm{erf}\left( \frac{\log_{10}\mvir -
        \log_{10}M_{\rm min}}{\sigma_{\log_{10}M}} \right) \right)
\end{equation}
For every host halo in the catalog we draw a random number from a uniform distribution $\mathcal{U}(0, 1);$ for a host halo of present-day virial mass $\mvir,$ a central galaxy is assigned to the halo if the associated random number is less than $\langle N_{\mathrm{cen}} \rangle(\mvir);$ halos with random values exceeding  $\langle N_{\mathrm{cen}} \rangle(\mvir)$ are left devoid of centrals.

We model the distribution $P(\nsat|\mvir)$ to be a Poisson distribution with first moment given by a power law, 
\begin{equation}
\label{eq:nsat}
\langle N_{\rm sat}\rangle(\mvir) = \left( \frac{\mvir - M_0}{M_1} \right)^{\alpha}. 
\end{equation}

@@

We model the velocities of satellite galaxies to be a Gaussian whose first moment is the host halo velocity and whose second moment is the solution to the isotropic Jeans equation for an NFW profile \citep{more09b}, 
\begin{equation}
\sigma^{2}_{r}(\tilde{r}|c) = V_{\rm vir}^{2}\frac{c^{2}\tilde{r}(1 + c\tilde{r})^{2}}{g(c)}\int_{c\tilde{r}}^{\infty}{\rm d}y\frac{g(y)}{y^{3}(1 + y)^{2}}, 
\end{equation}
where $\tilde{r} = r/\rvir,$ $g(x) = \rm{ln}(1+x) - x / (1+x),$ and $V_{\rm vir}^{2} = G\mvir/\rvir.$

@@

In addition to the standard occupation statistics described in the
previous section, in this paper we also use the decorated HOD
formalism to connect galaxies to dark matter halos in a manner that
has simultaneous dependence on both $\mvir$ and halo
concentration. Briefly, we use Equations \ref{eq:ncen} and
\ref{eq:nsat} as our ``baseline" first occupation moments. At fixed
$\mvir,$ halos are divided into one of two categories, those of high-
and low-concentration, depending on whether the concentration of the
halo places it above or below the rank-order percentile $f_{\rm
  split},$ which we keep fixed to $f_{\rm split}=0.5$ throughout the
paper. High-concentration halos have a different first occupation
moment relative to low-concentration halos of the same mass,
$\langle\ngal|\mvir, c_{\rm high}\rangle \neq \langle\ngal|\mvir,
c_{\rm low}\rangle.$ The difference between the first moment of high-
and low-concentration halos is modulated by $\abias,$ the novel
parameter of the decorated HOD governing assembly bias. Values of
$\abias=\pm1$ correspond to the maximum strength of assembly bias
allowable by the constraint that the model preserves the marginalized
first moment, $\langle\ngal|\mvir\rangle;$ thus regardless of the
value of $\abias,$ in the decorated HOD the marginalized first moment
of centrals and satellites are unchanged from the baseline value
defined by Equations \ref{eq:ncen} and \ref{eq:nsat}. We refer the
reader to \citet{hearin_etal16} for further details about the
decorated HOD.

@@

property that we assume to determine the galaxy HOD is halo formation time. $A_{\rm cen,sat}=0$ means that there is no 
assembly bias while $A_{\rm cen,sat}=1$ ($A_{\rm cen,sat}=-1$) means that halo formation time is maximally 
correlated (anticorrelated) with halo formation time. Thus the parameters range over $-1 \le A_{\rm cen,sat} \le 1$. 
If the constraints on $A_{\rm cen}$ and $A_{\rm sat}$ are unspecified, then the model used to interpret the data 

@@

A second noteworthy difference between the present work and that of
\citet{zehavi_etal11} is that we find many parameters to be notably
more poorly constrained. At the lower luminosity thresholds, for
example, we constrain $\log (M_{\rm min})$ and $\sigma_{\log M}$ with
several times lower precision than \citet{zehavi_etal11}. We do not
show our constraints on $\log (M_0)$ as they are very poor, with
1-sigma constraints $\gta 1$~dex for all samples. In several cases,
the constraint on $\log (M_0)$ is determined by the prior given in
Table~\ref{table:priors}. This is in stark contrast to several of the
results of \citet{zehavi_etal11}. For example, for the threshold
sample with $M_r < -19.5$ ($M_r < -20.5$), \citet{zehavi_etal11} quote
$\log (M_0) = 12.23 \pm 0.17$ ($12.35 \pm 0.24$), whereas we infer
$\log (M_0) = 11.38^{+0.95}_{-1.57}$
($11.19^{+0.89}_{-1.39}$). Examining the form of Eq.~(\ref{eq:nsat}),
it is sensible that the parameter $\log (M_0)$ should be unconstrained
at the lower end, because the value of $M_0$ does not alter the
predicted satellite number once $M_0 \ll M_1$. Therefore, it seems
likely that the tighter constraints quoted by \citet{zehavi_etal11}
must be an error.

We have confirmed with a subset of the authors of \citet{zehavi_etal11} 
that the number of MCMC samples they included in their analysis was 
insufficient in a number of cases and that this can lead to a significant underestimation of the uncertainties 
on the inferred parameters, especially $\log (M_{\rm min})$ and $\sigma_{\log M}$ 
(Z. Zheng \& I. Zehavi, Private Communication). The analysis of \citet{zehavi_etal11} 
used $10^4$ samples, whereas we find several $\times 10^6$ samples are often necessary for convergence. 
Additionally, we have recreated qualitatively similar behavior by considering only small subsets of our 
full MCMC chains. Consequently, insufficient sampling of the posterior seems to be the likely resolution 
of the discrepancies between our work and that of \citet{zehavi_etal11}.

Two degeneracies are manifest in Fig.~\ref{fig:Mr20triangle} that are common to all of our 
analyses. The parameters $\log (M_1)$ and $\alpha$ are degenerate with each other and positively 
correlated. The parameter $M_1$ is the mass scale at which a halo has one satellite on 
average and $\alpha$ is the power-law index describing the dependence of average satellite 
number on halo mass. Increasing $M_1$ decreases the number of satellites in large halos 
by placing satellites preferentially in halos of larger mass. An increase in $\alpha$ can partly 
compensate for an increase in $M_1$ by increasing the rate at which average satellite number 
grows with halo mass, leading to a larger number of satellites in high-mass halos.

@@
